Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-05-12 main.Rmd
revised the circular relativies section

The hierarchical structure of time creates a natural nested ordering which can produce multiple-order-up granularities. We shall use the notion of a hierarchy table to define them.

\begin{definition}\label{def:hierarchy} 
The {\bf hierarchy table} $H_n: (G, C, K)$ arranges $n$ periodic linear granularities from lowest to higher order and lays down the way they relate through $K$ so that cyclic granularities $C$ could be defined relative to these linear granularities. Let $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$. Then $K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$, if $C_{G_{l}, G_{m}}$ is a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$, if $C_{G_{l}, G_{m}}$ is quasi-circular.
\end{definition}

\begin{example}
{\rm So far we have used examples of linear granularities from Gregorian calendar as it is the most widely used calendar. However, it is far from being the only one. The day is the basic unit of time underlying most calendars [@Reingold2001-kf]. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds. Nevertheless, for any calendar a hierarchy table can be defined. For example, in the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun. \autoref{tab:tab-mayan} represents the hierarchy table for the Mayan calendar.}
\end{example}

@@

@@

\noindent It might not be always possible to position an aperiodic linear granularity in a hierarchy table. Hence, we assume that the hierarchy table consists of periodic linear granularities only and that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.

\begin{definition}\label{def:norderup} 
Cyclic granularities which are nested within multiple levels of the hierarchy table are called {\bf multiple-order-up} granularities.
\end{definition}

A _single-order-up_ is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
 <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->

\begin{example}
{\rm For \autoref{tab:tab-mayan}, examples of multiple-order-up granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.}
\end{example}

>>>>

The hierarchical structure of time creates a natural nested ordering  which can be used in the computation of relative pairs of granularities. This is illustrated using two examples, a Mayan calendar, which is simple, in addition to the more complex Gregorian calendar.

\begin{definition}\label{def:hierarchy} 
The nested ordering of linear granularities can be organized into a {\bf hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order. It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$. can be defined relative to the linear granularities. Let $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$. Then $K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$, if $C_{G_{l}, G_{m}}$ is a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$, if $C_{G_{l}, G_{m}}$ is quasi-circular.
\end{definition}

\begin{example}
{\rm  \autoref{tab:tab-mayan} shows the hierarchy table for the Mayan calendar. In the Mayan calendar, one day is referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal (about a year); 1 katun = 20 tun (20 years) and 1 baktun = 20 katun. }
\end{example}

@@

 <!--So far we have used examples of linear granularities from Gregorian calendar as it is the most widely used calendar. However, it is far from being the only one.-->
\noindent It is interesting to note, that in the Mayan calendar, like most calendars, that day is the basic unit of time [@Reingold2001-kf]. The structuring of larger units,  weeks, months, years and cycle of years, though, varies substantially. An exception is the French revolutionary calendar, which divides each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds. Nevertheless, for any calendar a hierarchy table can be defined.

@@

Note that, it is not always possible to organize an aperiodic linear granularity in a hierarchy table. Hence, we assume that the hierarchy table consists of periodic linear granularities only and that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular.

\begin{definition}\label{def:norderup} 
The hierarchy table contains {\bf multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
A {\bf single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
\end{definition}

 <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->

\begin{example}
{\rm In \autoref{tab:tab-mayan}, kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.}
\end{example}
